---
output:
  md_document:
    variant: markdown_github
---

[![Build Status](https://travis-ci.org/rekonstrukt/swedishbutterflies.svg?branch=master)](https://travis-ci.org/rekonstrukt/swedishbutterflies)  

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

```{r include=FALSE}
knitr::opts_chunk$set(warning=FALSE, comment=NA)
```

`swedishbutterflies` is an R package for SeBMS - the Swedish Butterfly Monitoring Scheme - offering tools for accessing data, making plots and a Shiny app.

## Installing from github

If you want to install the latest version:

```{r message=FALSE, warning=FALSE, results="hide", eval=FALSE}
# First make sure you have the devtools package
# which simplifies installations from github
# Note: Windows users have to first install Rtools to use devtools

install.packages("devtools") 
library(devtools)
install_github("rekonstrukt/swedishbutterflies")

```

## Quick start

Load the package in your R environment and create a `config.yml` with the db connection details:

```{r message=FALSE, warning=FALSE, eval=FALSE}

library(swedishbutterflies)
library(rappdirs)

# location for config.yml with db connection details
app_dir("sebms")$config()
```

Example content that can be used in the `config.yml`:

```{r message=FALSE, warning=FALSE, echo=FALSE}

# example content in config.yml
content <- readLines(system.file("extdata", "config.yml", 
  package = "swedishbutterflies"))

cat(paste(content, collapse = "\n"))
```

Before being able to access data from the database, set up the `config.yml` and if you use environment variables in it for uid or password, also set up your `.Renviron` with that content:

```console
DBUSER = your_db_user
DBPASS = your_db_pwd

```

Then, to see some quick usage examples to get you started, read the Vignette.

## Credits

The package bundles code and data assembled and curated by Lars Petterson at https://dagfjarilar.lu.se

## Meta

* Please [report any issues or bugs](https://github.com/rekonstrukt/swedishbutterflies/issues).
* License: AGPL
