---
title: "Using swedishbutterflies - tools from SeBMS"
author: "Markus Skyttner"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{sebms-intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

## Using "swedishbutterflies"

The R-package "swedishbutterflies" contains functions for retrieving data from SeBMS - the Swedish Butterfly Monitoring Scheme's database. It can be used for accessing data, for generating static figures and for using an included Shiny proof-of-concept app.

### Retrieving data

Some data is bundled into the package, for example:

```{r, message=FALSE}

library(swedishbutterflies)

suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(knitr))

# first few rows of precipitation 2015 data
sebms_data_precip_2015 %>% 
head(5) %>% 
kable()

# same for temperature 2015 data
sebms_data_temp_2015 %>% 
head(5) %>% 
kable()

# species counts
sebms_data_specieslist_cum %>%
head(5) %>%
kable()

# species count histograms

sebms_data_species_histo %>%
head(5) %>%
kable()

sebms_data_species_per_site_sitetype %>%
head(5) %>%
kable()

# Naturum site locations
sebms_data_sites_naturum %>%
head(5) %>% 
kable()

```

There are also functions for retrieving data from the database:

```{r}

sebms_species_per_year() %>% 
head(5) %>% 
kable()

sebms_species_per_year_filtered() %>% 
head(5) %>% 
kable()

sebms_species_per_year_site_filtered() %>% 
head(5) %>% 
kable()

sebms_species_per_year_site_counts_filtered() %>%
head(5) %>% 
kable()
  
```

### Generating figures

The figures can be included in an annual report, for example this figure with temperature and precipitation data for 2015.

```{r, message=FALSE, fig.show='hold', fig.cap='Temperature and Precipitation for 2015', fig.width=7}

library(magick)

sebms_precip_temp_2015_plot() %>% 
  image_resize("700x")

```

Two plots for cumulative species counts can be generated like this:

```{r}
plots <- sebms_specieslist_cum_plots()

p1 <- plots$p1
p2 <- plots$p2
```

The first one looks like this:

```{r, fig.width=7, fig.height=7}
p1
```

And the second one looks like this:

```{r, fig.width=7, fig.height=9}
p2
```

Static versions can also be saved given pixel-based width and height:

```{r}

p1 %>% sebms_ggsave(661, 812, filename = "/tmp/specieslist-01.png")
p2 %>% sebms_ggsave(666, 900, filename = "/tmp/specieslist-02.png")

```

Species count histograms:

```{r, fig.width=7}
sebms_species_histo_plot()
```

```{r, fig.width=7}
sebms_species_per_site_sitetype_plot()
```

Sites per province and year:

```{r}

library(tibble)
library(DT)

sites <- sebms_data_sites_per_year_province

df <- 
  t(sites) %>% 
  as_tibble() %>% 
  setNames(sites$province)

datatable(df)

```

Show some climate data for Naturum sites where butterflies might currently be observed:

```{r}

gogogo <- 
  sebms_naturum_climate() %>%
  filter(is_gogogo) %>%
  select(id, is_gogogo, t, ws, tcc_mean, pmean, date)

gogogo %>% head(5) %>% kable

```

